---
title: "Statistical Inference Part 2"
author: "Gill Collier"
date: "07/11/2020"
output: pdf_document
---

# Gill Collier - 01 November 2020
# Coursera: Statistical Inference - Course Project

# Part 2: Basic Inferential Data Analysis
# Now in the second portion of the project, we're going to analyze the 
# ToothGrowth data in the R datasets package.

# Libraries
```{r}
library(ggplot2)
```

# Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
data = ToothGrowth
?"ToothGrowth"
```

# The ToothGrowth data shows the Effect of Vitamin C on Tooth Growth in Guinea Pigs
# Description
# The response is the length of odontoblasts (cells responsible for tooth 
# growth) in 60 guinea pigs. Each animal received one of three dose levels of 
# vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange 
# juice or ascorbic acid (a form of vitamin C and coded as VC).

```{r}
str(ToothGrowth)
head(ToothGrowth)
```
# Summarise the data
```{r}
summary(ToothGrowth)

ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
        geom_bar(stat="identity") +
        facet_grid(. ~ supp) +
        ggtitle("Tooth length by Vitamin C dosage for delivery methods") +
        xlab("Vitamin C Dosage") +
        ylab("Tooth length")
```

# Use confidence intervals and/or hypothesis tests to compare tooth growth 
# by supp and dose.
```{r}
t.test(len ~ supp, data = ToothGrowth)
t.test(len ~ supp, data = subset(ToothGrowth, dose == 0.5))
t.test(len ~ supp, data = subset(ToothGrowth, dose == 1))
t.test(len ~ supp, data = subset(ToothGrowth, dose == 2))
```

# State your conclusions and the assumptions needed for your conclusions.
# Tooth Length does increase as the dosage is increase for both delivery methods

# When the lower dosages are administered (0.5 and 1) the most effective
# method seems to the OJ but this is not the case with the higher dosage
# The t-test analysis shows that the delivery method has no real effect 
# on tooth growth.