---
title: "Coursera Statistical Inference Course Project part 1"
author: "Gill Collier"
date: "06/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Gill Collier - 01 November 2020
# Coursera: Statistical Inference - Course Project

# Part 1: Simulation Exercise Instructions
# In this project you will investigate the exponential distribution in R and 
# compare it with the Central Limit Theorem. The exponential distribution can 
# be simulated in R with rexp(n, lambda) where lambda is the rate parameter. 
# The mean of exponential distribution is 1/lambda and the standard deviation 
# is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will 
# investigate the distribution of averages of 40 exponentials. Note that you 
# will need to do a thousand simulations.

# Illustrate via simulation and associated explanatory text the properties of 
# the distribution of the mean of 40 exponentials. You should:
# 1. Show the sample mean and compare it to the theoretical mean of the 
# distribution.
# 2. Show how variable the sample is (via variance) and compare it to the 
# theoretical variance of the distribution.
# 3. Show that the distribution is approximately normal.
# In point 3, focus on the difference between the distribution of a large 
# collection of random exponentials and the distribution of a large collection 
# of averages of 40 exponentials.

# Libraries
```{r}
library(ggplot2)
```

# Set variables
```{r}
set.seed(123)
lambda <- 0.2
n <- 40
```

# Generate the sample means
```{r}
mns = NULL
for (i in 1 : 1500) mns = c(mns, mean(rexp(n, lambda)))
summary(mns)
```

# Calculate the mean of these means
```{r}
m_mns <- mean(mns)
```


# Calculate the theoretical mean
```{r}
t_mns <- lambda^-1
t_mns
abs(m_mns - t_mns)
```

# The central limit theorem states that the sampling distribution of a sample 
# mean is approximately normal if the sample size is large enough, even if 
# the population distribution is not normal. 

# The CLT in this simulation appears to be valid as this shows that increasing 
# the number of samples narrows the gap between the simulation mean and the 
# theoretical mean

# Calculate the sample variance
```{r}
s_var <- var(mns)
s_var
```

# Calculate the theoretical variance
```{r}
t_var <- (lambda * sqrt(n)) ^ -2
t_var
```

# Compare the sample variance to the theoretical variance
```{r}
s_var - t_var
```

# This comparison shows there is only a small difference between the 
# simulation variance and the theorectical variance

# The distribution can be shown in a histogram. This shows the sample means 
# from the simulation with an overlay of a normal distribution.
```{r}
hist(mns, prob = TRUE, col = "light blue", main = 
             "Histogram of Means from Simulation", xlab = "simulation mean", 
     breaks = 20)
```
